Data Science Applied to Ag - Final Project - ML

### 2025-04-26 code/Draft_script_No_Cat_Features.qmd
d

weather_test <- site_year_test %>%
  mutate(weather = pmap(
    list(.y = year, .site = site, .lat = lat, .lon = lon),
    ~ download_daymet(site = .site,
                      lat   = .lat,
                      lon   = .lon,
                      start = .y,
                      end   = .y,
                      simplify = TRUE,
                      silent = TRUE)
  )) %>%
  select(year, site, weather) %>%
  unnest(weather) %>%
  pivot_wider(names_from = measurement, values_from = value) %>%
  clean_names()

@@

# Feature engineering monthly summaries for 2024 site‐years
fe_test <- test_full %>%
  # create date & month
  mutate(date = as.Date(paste0(year, "/", yday), "%Y/%j"),
         month_abb = month(date, label = TRUE)) %>%
  # select and rename to match training
  transmute(
    year, site, hybrid,
    month_abb,
    soil.ph       = soilp_h,    # adjust these names if your clean_names() differ
    soil.om.pct   = om_pct,
    soil.k.ppm    = soilk_ppm,
    soil.p.ppm    = soilp_ppm,
    dayl.s        = dayl_s,
    prcp.mm       = prcp_mm_day,
    srad.wm2      = srad_w_m_2,
    tmax.c        = tmax_deg_c,
    tmin.c        = tmin_deg_c,
    vp.pa         = vp_pa
  ) %>%
  group_by(year, site, hybrid, month_abb) %>%
  summarise(
    mean_soil.ph     = mean(soil.ph,   na.rm = TRUE),
    mean_soil.om.pct = mean(soil.om.pct, na.rm = TRUE),
    mean_soil.k.ppm  = mean(soil.k.ppm,  na.rm = TRUE),
    mean_soil.p.ppm  = mean(soil.p.ppm,  na.rm = TRUE),
    mean_dayl.s      = mean(dayl.s,      na.rm = TRUE),
    sum_prcp.mm      = sum(prcp.mm,      na.rm = TRUE),
    mean_srad.wm2    = mean(srad.wm2,    na.rm = TRUE),
    mean_tmax.c      = mean(tmax.c,      na.rm = TRUE),
    mean_tmin.c      = mean(tmin.c,      na.rm = TRUE),
    mean_vp.pa       = mean(vp.pa,       na.omit = TRUE)
  ) %>%
  ungroup() %>%
  pivot_longer(
    cols = starts_with(c("mean_", "sum_")),
    names_to = "name",
    values_to = "value"
  ) %>%
  mutate(varname = paste0(name, "_", month_abb)) %>%
  select(-name, -month_abb) %>%
  pivot_wider(names_from = varname, values_from = value) %>%
  select(-ends_with("_NA"))

>>>>

weather_test <- site_year_test %>%
  mutate(weather = pmap(
    list(.y = year, .site = site, .lat = lat, .lon = lon),
    ~ download_daymet(site = .site,
                      lat   = .lat,
                      lon   = .lon,
                      start = .y,
                      end   = .y,
                      simplify = TRUE,
                      silent = TRUE)
  )) %>%
  select(year, site, weather) %>%
  unnest(weather) %>%
  pivot_wider(names_from = measurement, values_from = value) %>%
  #clean_names()

@@

# Feature engineering monthly summaries for 2024 site‐years
fe_test <- test_full %>%
  # creating date & month
  mutate(date = as.Date(paste0(year, "/", yday), "%Y/%j"),
         month_abb = month(date, label = TRUE)) %>%
  # selecting and renaming to match training
  transmute(
    year, site, hybrid,
    month_abb,
    soil.ph       = soilp_h,    # will have to adjust these names if the clean_names() differ
    soil.om.pct   = om_pct,
    soil.k.ppm    = soilk_ppm,
    soil.p.ppm    = soilp_ppm,
    dayl.s        = dayl_s,
    prcp.mm       = prcp_mm_day,
    srad.wm2      = srad_w_m_2,
    tmax.c        = tmax_deg_c,
    tmin.c        = tmin_deg_c,
    vp.pa         = vp_pa
  ) %>%
  group_by(year, site, hybrid, month_abb) %>%
  summarise(
    mean_soil.ph     = mean(soil.ph,   na.rm = TRUE),
    mean_soil.om.pct = mean(soil.om.pct, na.rm = TRUE),
    mean_soil.k.ppm  = mean(soil.k.ppm,  na.rm = TRUE),
    mean_soil.p.ppm  = mean(soil.p.ppm,  na.rm = TRUE),
    mean_dayl.s      = mean(dayl.s,      na.rm = TRUE),
    sum_prcp.mm      = sum(prcp.mm,      na.rm = TRUE),
    mean_srad.wm2    = mean(srad.wm2,    na.rm = TRUE),
    mean_tmax.c      = mean(tmax.c,      na.rm = TRUE),
    mean_tmin.c      = mean(tmin.c,      na.rm = TRUE),
    mean_vp.pa       = mean(vp.pa,       na.omit = TRUE)
  ) %>%
  ungroup() %>%
  pivot_longer(
    cols = starts_with(c("mean_", "sum_")),
    names_to = "name",
    values_to = "value"
  ) %>%
  mutate(varname = paste0(name, "_", month_abb)) %>%
  select(-name, -month_abb) %>%
  pivot_wider(names_from = varname, values_from = value) %>%
  select(-ends_with("_NA"))
